Workplan

### 2018-05-14 LiteratureReview/LiteratureReview.tex
updated LR and QR
updated

\par
According to a recent report commissioned by the Alzheimer's Society, they estimate the prevalence of Dementia in the UK at approximately 815,000 people. This represents 1 in 14 of the population aged 65 or over. This report also estimates an annual healthcare spend on £4.3 billion of which approximately £85 million is spend solely on diagnosis. They also estimate that the overall impact of dementia (excluding the costs associated with early onset dementia) is £26.3 billion annually. Globally, this picture is a lot bleaker. A recent report suggests that in 2015 there were 46 million people with a diagnosis of dementia and that number is expected to hit 131.5 million by 2050 \cite{Prince2015}. The report also states that the worldwide cost of dementia in 2018 is estimated to be in the region of one trillon US dollars.
\newline
\par
The purpose of this review is to seek to understand how dementia impacts speech and language production from a psychological and linguistic perspective with a view to potentially using natural language processing and technology to aid in the early diagnosis of dementia and/or Mild Cognitive Impairment.
\newline
\par
A search of the literature was conducted mainly from a psychological perspective as this would inform the theoretical side of things. Next, a search of the literature from a Computer Science perspective was also made with a view to exploring the how domain of machine learning and more specifically natural language processing can be applied to this problem. Finally a search was carried out to see what work has already been done in this area.

@@

\section{Current ways of diagnosing Dementia using Language}
\subsection{Free Cued Selective Recall Test}
The Free Cued Selective Recall Test (FCSRT) was created by Grober and his team \cite{Grober1988}.

@@

>>>>

\par
According to a recent report commissioned by the Alzheimer's Society, they estimate the prevalence of Dementia in the UK at approximately 815,000 people. This represents 1 in 14 of the population aged 65 or over. This report also estimates an annual healthcare spend on £4.3 billion of which approximately £85 million is spend solely on diagnosis. They also estimate that the overall impact of dementia (excluding the costs associated with early onset dementia) is £26.3 billion annually. Globally, this picture is a lot bleaker. A recent report suggests that in 2015 there were 46 million people with a diagnosis of dementia and that number is expected to hit 131.5 million by 2050 \cite{Prince2015}. The report also states that the worldwide cost of dementia in 2018 is estimated to be in the region of one trillon US dollars.
\newline
\par
The purpose of this review is to seek to understand how dementia impacts speech and language production from a psychological and linguistic perspective with a view to potentially using natural language processing and technology to aid in the early diagnosis of dementia and/or Mild Cognitive Impairment.
\newline
\par 
Aims and Objectives:
\newline
\par
A search of the literature was conducted mainly from a psychological perspective as this would inform the theoretical side of things. Next, a search of the literature from a Computer Science perspective was also made with a view to exploring the how domain of machine learning and more specifically natural language processing can be applied to this problem. Finally a search was carried out to see what work has already been done in this area.

@@

\section{Current ways of diagnosing Dementia using Language}
\subsection{Free Cued Selective Recall Test}
The Free Cued Selective Recall Test (FCSRT) was created by Grober and his team \cite{Grober1988}. This is a test that primarily tests memory but also tests a subjects ability to name objects. Four items are presented at a time on a card and each participant is asked to name and recall with and without cues.

@@

\section{stuff to add}
Annotated Bibliography for the project "Ambient Monitoring of Speech for Evidence of Cognitive Decline".

Orimaye, S. O., Wong, J. S-M., Golden, K. J., Wong, C. P., & Soyiri, I. N. (2017) Predicting probably Alzheimer's Disease using linguistic deficits and biomarkers.

There is an increasing body of literature which supports the use of machine learning to analyse speech to evidence cognitive decline. Orimaye et al (2017) investigate the use of machine learning algorithms to detect differences in syntactic, lexical and n-gram linguistic biomarkers to distinguish between those with probably AD and healthy controls, The authors found significant differences in the uses of all three types of biomarkers for those with AD and healthy controls. Their results show that the top 1000 n-gram features plus the twenty-three syntactic and lexical features was the most successful at differentiating the two groups (0.93 AUC)). However, there were some limitations of the study. The context of the audio content was very specific, limited to one specific description task, and therefore using general speech as proposed in this project would potentially use a different set of linguistic features which may or may not have similar predictive power. In addition, the transcriptions were encoded using the CHAT format which is a framework for manually annotating speech and the challenge would be finding a way to automate this. This paper is noteworthy as it provides evidence that using machine learning to identify participants with MCI is possible and it provides some ideas of potential features which can be used. Unique to this paper is the use of n-gram features, which could be explored in the proposed project.

Asgari, M., Kaye, J., & Dodge, H. (2017) Predicting mild cognitive impairment from spontaneous spoken utterances. Alzheimers & Dementia: Translational Research & Clinical Interventions 3, 219 - 228

Asgari, Kaye and Dodge (2017) also looked at the linguistic characteristics of older adults with mild cognitive impairment (MCI) vs healthy controls where they hypothesised that they would be able to predict those with MCI, a distinguishing characteristic of Alzheimer's disease and other variations of dementia. Using recordings of unstructured conversations (with standardized preselected topics across subjects) between interviewers and interviewees they grouped spoken words using Linguistic Inquiry and Word Count (LIWC) which is a technique used to categorize words into features such as negative and positive words. They then applied support vector machines (SVM's) and random forest classification algorithms to investigate whether machine learning could be used to distinguish between those with MCI and healthy controls. They were able to successfully used machine learning algorithms to distinguish between these two groups with an accuracy of 84%. The authors report that this method is highly reliant on high-fidelity transcription of the conversations which is labour intensive, but they anticipate that technology is advancing sufficiently quickly that automated high-quality transcriptions are possible in the near future. This paper provides a different perspective in how to tackle the problem deriving language features from speech using a different framework. The challenges both Orimaye et al and these authors experienced were around the automation of transcription and this would potentially be an area to explore in the proposed project.

Pennebaker, J.W., Boyd, R. L., Jordan, K and Blackburn, K., (2015) The development and psychometric properties of LIWC2015

Asgari et al (2017) used the Linguistic Inquiry and Word Count (LIWC) methodology, developed by Pennebaker et al (2007) to generate features for their paper. Pennebaker, Boyd, Jordan and Blackburn (2015) have since developed their framework. The LIWC is designed to categorize and evaluate the various emotional, cognitive and structural components which are present within samples of speech and/or written text. This method, Linquistic Inquiry & Word Count (LIWC) has evolved from it's initial incarnation in 1993 to it's last update in 2015. The premise is the use of words from particular categories provide an insight into the psychological processes and/or diagnoses an individual has. The latest version of this methodology uses a dictionary of 6400 words, word stems and select emoticons and assigns these constructs to various categories. This has been updated to include modern uses of language such as 'text speak'. This manual goes summarises the process of constructing the framework as well as discussing the reliability and validity of the measure and a review of studies using the LIWC was conducted by Tausczik and Pennebaker (2010) support the notion that the LIWC is valid across multiple psychological domains. This paper is useful as it provides a useful and validated framework from which the proposed project can potentially derive language features.

Schwartz, A. H., Eichstaedt, J., Kern, M. L., Park, G., Sap. M., Stillwell. D., . . . Ungar. L. (2014)  Towards Assessing Changes in Degree of Depression through Facebook. Workshop on Computational Linguistics and Clinical Psychology: From Linguistic Signal to Clinical Reality, 118-125 
Currently formal diagnosis of depression is expensive and time-consuming involving both the use of questionnaires and the use of a trained professional to assess an individual. Schwartz et al, explored the use of language as an aid to diagnose depression in a naturalistic setting i.e. facebook status updates. The authors used both the LIWC (described above) and n-grams as features to distinguish those with depression from healthy individuals and were able to track participants levels' of depression through language successfully.Whilst the authors were able to use a reasonably sound method of measure degree of depression and have evidenced a methodology that is able to track a levels of depression over time, it would be more helpful to use a more clinically relevant measure of depression and anxiety i.e., PHQ-9 (Kroenke, 2001) and GAD-7 (Spitzer et al, 2006) to add validity to their findings. 
This paper is interesting as it uses both LIWC and n-grams as a method of deriving features from text / speech for depression which, as described above, have also been used as features to distinguish MCI from healthy participants. Given the comorbity of depression and dementia (Meyers, 1998), it would be interesting for the proposed project to further analyse the link between dementia and clinical depression via the use of language. This paper also supports the idea that language changes over time and is a marker for deterioration in both those with depression and dementia.

### 2018-05-14 QualifyingReport/QualifyingReport.tex
updated

\begin{document}
	\maketitle
	\newpage
	\section{Overview of Problem}
	Dementia has been identified as one of those fast growing difficulties facing the world. A recent report suggests that in 2015 there were 46 million people with a diagnosis of dementia and that number is expected to hit 131.5 million by 2050(CITE). The report also states that the worldwide cost of dementia in 2018 is estimated to be in the region of one trillion US dollars.
	\newline
	\par
	In 2009, the Department of Health in it's National Dementia strategy made 'early diagnosis and support' one of it's key themes as part of this strategy. A lot of work, therefore, has gone into trying to find ways of improving the early diagnosis of Alzheimer's Disease (AD) and Mild Cogntive Impairment (MCI). As described above, the numbers of those suffering from AD and MCI are going to increase as the population ages and thus it is important that we utilise technology wherever possible to aid clinicians. At the present time diagnosis is typically conducted at memory clinics by trained clinicians. I theorise that we might be able to use diagnose patients in their own homes utilising smart-home technology. 
	\newline
	\par
	Language analysis is an important part of assessing those suspected of suffering with AD or MCI and this language analysis in part informs an official clinical diagnosis. This analysis often takes the form of tasks such as Free Cued Selective Recall Task, the Boston Cognitive Assessment, and the Mini Mental State Examination which are all quite targeted in the language required of the person being assessed. 
	\newline
	\par
	A step towards the ambient detection of AD and/or MCI in population requires that we look for similar features in language used naturally as part of every day conversation. Part of this work therefore is to see whether spontaneous discourse, such a semi-structured interview is a task that has the ability to put pressure both the cognitive and linguistic systems as it requires the ability to understand speech, formulating a coherent answer and then responding in an appropriate way. Berisha \cite{Berisha2015}, has shown through language analysis in this area that there are marked differences in this process between those who have a diagnosis of AD and healthy controls. The next part of this work looks at how technology, particularly the use of smart home devices such as the Amazon Echo and Google Home families of devices can be used to capture data that can then be analysed.
	\newline
	\par
	The potential impact of this research is immense. Research has shown that early diagnosis of people with AD or MCI improves sufferers quality of life and can slow the progress of the disease. Equally, early diagnosis can increase the number of research opportunities for understanding the early stages of dementia and how the disease progresses so that more research can be conducted which may, in the future, lead to new treatments and other interventions.

@@

>>>>

\begin{document}
	\maketitle
	\newpage
	\section{Overview of Problem}
	Dementia has been identified as one of those fast growing difficulties facing the world. A recent report suggests that in 2015 there were 46 million people with a diagnosis of dementia and that number is expected to hit 131.5 million by 2050(CITE). The report also states that the worldwide cost of dementia in 2018 is estimated to be in the region of one trillion US dollars.
	\newline
	\par
	In 2009, the Department of Health in it's National Dementia strategy made 'early diagnosis and support' one of it's key themes as part of this strategy. A lot of work, therefore, has gone into trying to find ways of improving the early diagnosis of Alzheimer's Disease (AD) and Mild Cogntive Impairment (MCI). As described above, the numbers of those suffering from AD and MCI are going to increase as the population ages and thus it is important that we utilise technology wherever possible to aid clinicians. At the present time diagnosis is typically conducted at memory clinics by trained clinicians. I theorise that we might be able to use diagnose patients in their own homes utilising smart-home technology. 
	\newline
	\par
	Language analysis is an important part of assessing those suspected of suffering with AD or MCI and this language analysis in part informs an official clinical diagnosis. This analysis often takes the form of tasks such as Free Cued Selective Recall Task, the Boston Cognitive Assessment, and the Mini Mental State Examination which are all quite targeted in the language required of the person being assessed. 
	\newline
	\par
	A step towards the ambient detection of AD and/or MCI in population requires that we look for similar features in language used naturally as part of every day conversation. Part of this work therefore is to see whether spontaneous discourse, such a semi-structured interview is a task that has the ability to put pressure both the cognitive and linguistic systems as it requires the ability to understand speech, formulating a coherent answer and then responding in an appropriate way. Berisha et al \cite{Berisha2015}, has shown through language analysis in this area that there are marked differences in this process between those who have a diagnosis of AD and healthy controls. The next part of this work looks at how technology, particularly the use of smart home devices such as the Amazon Echo and Google Home families of devices can be used to capture data that can then be analysed.
	\newline
	\par
	The potential impact of this research is immense. Research has shown that early diagnosis of people with AD or MCI improves sufferers quality of life and can slow the progress of the disease. Equally, early diagnosis can increase the number of research opportunities for understanding the early stages of dementia and how the disease progresses so that more research can be conducted which may, in the future, lead to new treatments and other interventions.

@@

	%Complete appendices for each of the assessment. How do you do appendices in LaTeX

### 2018-05-14 Workplan/Workplan.tex
updated Lit Review
changes to Literature Review and QR
updated

	\section{Proposed Timeline, Key Dates and Deliverables}
	\subsection{April 2018}
	Thursday, 26th April - Workplan (First Draft Agreed)
	\subsection{May 2018}
	Wednesday, 9th May - Alexa Skills Training (All Day)
	\subsection{June 2018}
	Friday 15th June - First Draft of Literature Review to be handed in.
	\newline
	Monday 18th June - Monday 25th June - Annual Leave
	\subsection{July 2018}

@@

	\subsection{January 2019}
	Friday 11th January - Rough Draft of Qualifying Due
	Friday 18th January - Complete first rough pass of capturing and processing speech.

	\subsection{February 2019}
	Friday, 1st February - Qualifying Report Due

	\subsection{March 2018}

	\subsection{April 2018}
	Monday, 1st April - Viva Due by this point.
\end{document}

>>>>

	\section{Proposed Timeline, Key Dates and Deliverables}
	\subsection{April 2018}
	Thursday, 26th April - Workplan (First Draft Agreed)
	\subsection{May 2018}
	Wednesday, 9th May - Alexa Skills Training (All Day)
	\subsection{June 2018}
	Friday 15th June - First Draft of Literature Review to be handed in.
	\newline
	Monday 18th June - Monday 25th June - Annual Leave
	\subsection{July 2018}
	Friday 13th July - Written Assignment (IS4001) Due
	Friday 13th July - Project Plan (IS4001) Due

@@

	\subsection{January 2019}
	Friday 11th January - Rough Draft of Qualifying Due
	Friday 18th January - Complete first rough pass of capturing and processing speech.
	Thursday, 31st January - Qualifying Report Due

	\subsection{February 2019}

	\subsection{March 2019}

	\subsection{April 2019}
	Monday, 1st April - Viva Due by this point.
\end{document}
